---
title: "Week 7 · Class 1: Practice Exercises"
subtitle: "Multivariate Models & Categorical Predictors and Interactions"
date: "2026"
format:
  pdf:
    toc: false
    number-sections: true
    colorlinks: true
---

```{r}
#| include: false
library(tidyverse)
library(knitr)
library(formatR)
# When running chunks interactively, set the working directory to this file's
# folder so the relative data/ paths resolve. (During render, knitr already
# uses the document's directory, so this is skipped.)
if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable()) {
  .doc_path <- rstudioapi::getSourceEditorContext()$path
  if (length(.doc_path) == 1 && nzchar(.doc_path)) setwd(dirname(.doc_path))
}
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = FALSE)
```

# Segment A — Multivariate Models

# Non-AI Exercises

## 1. Understanding Multivariate Regression

### 1.1 Multiple Choice: Why Control Variables?

Why do we include control variables in regression models?

a) To make the model more complex
b) To isolate the effect of our variable of interest
c) To increase R-squared
d) To reduce sample size

Answer: _____

### 1.2 Fill in the Blanks: Multiple Regression

In the equation Y = β₀ + β₁X₁ + β₂X₂ + ε:

1. β₁ represents the effect of X₁ holding _____ constant
2. This is called the _____ effect of X₁
3. β₀ is the predicted Y when both X₁ and X₂ equal _____
4. Adding X₂ helps control for _____ variables
5. The model assumes X₁ and X₂ have _____ effects

Word bank: X₂, partial, zero, confounding, additive

### 1.3 True or False: Multivariate Models

Mark each statement as True (T) or False (F):

_____ Control variables must be correlated with the outcome  
_____ Adding controls always improves causal inference  
_____ Multivariate models can have multiple independent variables   
_____ We should control for every available variable  

## 2. Simpson's Paradox

### 2.1 Match: Simpson's Paradox Examples

Match each scenario with what it demonstrates:

**Scenarios:**

a) Overall correlation reverses within subgroups
b) Aggregated data hides important patterns
c) A third variable explains the relationship
d) Different groups have different relationships

**Concepts:**

1. Simpson's Paradox
2. Confounding
3. Heterogeneous effects
4. Aggregation bias

Matches: a = _____, b = _____, c = _____, d = _____

### 2.2 Code Detective: Controlling

What does this regression control for?

```r
model <- lm(conflicts ~ military_spending + regime_type + gdp_per_capita, 
            data = country_data)
```

This model controls for: _____

### 2.3 Multiple Choice: Interpretation

In a model with Y = 50 + 2X₁ - 3X₂, what does the coefficient -3 mean?

a) X₂ has a negative correlation with Y
b) Holding X₁ constant, a one-unit increase in X₂ decreases Y by 3
c) X₂ is three times less important than X₁
d) The model fit is poor

Answer: _____

## 3. Model Building

### 3.1 Fill in the Code: Adding Controls

Complete this code to build models progressively:

```r
# Simple model
model1 <- lm(outcome ~ treatment, data = data)

# Add one control (age)
model2 <- lm(outcome ~ treatment + _____, data = data)

# Add multiple controls (age + sex)
model3 <- lm(outcome ~ treatment + _____ + _____, data = data)

```

### 3.2 Match: Variable Selection

Match each criterion with its purpose:

**Criteria:**

a) Theory-driven selection
b) Statistical significance
c) Change in main coefficient
d) Model fit (R²)

**Purposes:**

1. See if controls affect our main relationship
2. Choose variables based on prior knowledge
3. Maximize explained variance
4. Include only "significant" predictors

Matches: a = _____, b = _____, c = _____, d = _____

### 3.3 Spot the Error: Bad Controls

What's wrong with this model?

```r
# Predicting election outcome
lm(won_election ~ campaign_spending + votes_received, data = election_data)
```

Problem: _____

## 4. Interpreting Multivariate Results

### 4.1 True or False: Coefficient Changes

When adding a control variable:

_____ The main coefficient always decreases  
_____ A large change suggests confounding  
_____ No change means the control is unimportant  
_____ The sign can flip from positive to negative  
_____ Standard errors always increase

### 4.2 Multiple Choice: Partial Effects

"Holding all else constant" means:

a) Other variables don't matter
b) We're comparing units with the same values on control variables
c) The controls are fixed at their means
d) There's no variation in control variables

Answer: _____

### 4.3 Multiple Choice: R² vs. Adjusted R²

Model A uses 3 predictors: R² = 0.42, adjusted R² = 0.40. Model B adds 5 more predictors: R² = 0.45, but adjusted R² falls to 0.38. Which conclusion is best?

a) Model B fits better — always prefer the higher R²
b) The extra predictors add little beyond chance; adjusted R² penalizes them, so prefer Model A
c) Both metrics agree that Model B is better
d) Adjusted R² is unreliable whenever it disagrees with R²

Answer: _____

# AI Exercises

**Tips for Working with Claude:**

- Ask for **R code using only tidyverse** (no other packages)
- Request **simple, focused answers** to your specific question—not complex analyses
- Ask Claude to **explain what the code is doing** since you're learning
- Avoid asking for visualizations or plots in these exercises
- Include the output of `glimpse()` in your prompt so Claude knows your variable names

**Example prompt:** "Using tidyverse in R, run a multiple regression with lm() predicting conflicts from trade_volume while controlling for distance_km and both_democracy. Explain how adding controls changes the trade_volume coefficient. Keep the code simple and explain what each line does. Here is what my data looks like: [paste glimpse output]"

For each AI exercise:
- Work with Claude to analyze the data
- Record your prompts and key findings

## 5. Trade and Conflict Analysis

**Dataset: data/trade_conflict_data.csv**

**Description**: Country-pair data on trade relationships and military conflicts.

**Variables**:

- `country_pair`: Country dyad identifier (chr)
- `year`: Year of observation (int)
- `trade_volume`: Bilateral trade in billions (dbl)
- `conflicts`: Number of militarized disputes (int)
- `distance_km`: Geographic distance (dbl)
- `both_democracy`: Both countries democratic (lgl)
- `alliance`: Military alliance exists (lgl)
- `gdp_sum`: Combined GDP in trillions (dbl)
- `colonial_history`: Former colonial relationship (lgl)
- `common_language`: Share official language (lgl)

### 5.1 Simple vs Controlled Analysis

```{r}
# Load the dataset
trade_data <- read_csv("data/trade_conflict_data.csv")

# Examine the structure
glimpse(trade_data)
```

Work with Claude to compare the `trade_volume`-`conflicts` relationship with and without controls.

### 5.2 Building Models Progressively

Ask Claude to help you build a model of `conflicts` predicted by `trade_volume`:

- Start with `trade_volume` → `conflicts`
- Add geographic/cultural controls (`distance_km`, `common_language`, `colonial_history`)
- Add political controls (`both_democracy`, `alliance`)
- Interpret how the `trade_volume` coefficient changes

```{r}

```

### 5.3 Testing Mechanisms

Work with Claude to explore whether `both_democracy` or `alliance` moderate the `trade_volume`-`conflicts` relationship by including interaction terms.

```{r}

```

## 6. Education and Income

**Dataset: data/education_income_panel.csv**

**Description**: Individual-level data on education, income, and background.

**Variables**:

- `person_id`: Individual identifier (int)
- `income`: Annual income in thousands (dbl)
- `years_education`: Years of schooling (int)
- `age`: Age in years (int)
- `experience`: Years of work experience (int)
- `parent_education`: Parents' average education (dbl)
- `ability_score`: Standardized test score (dbl)
- `urban`: Lives in urban area (lgl)
- `gender`: Gender (chr)
- `race`: Race/ethnicity (chr)

### 6.1 Returns to Education

```{r}
# Load the dataset
education_data <- read_csv("data/education_income_panel.csv")

# Check the structure
glimpse(education_data)
```

Use Claude to estimate the returns to `years_education` on `income` with increasing controls (`age`, `experience`, `urban`, `gender`, `race`).

### 6.2 Omitted Variable Bias

Use Claude to estimate returns to `years_education` on `income` with increasing controls.

Work with Claude to:
- Show how excluding `ability_score` biases the `years_education` coefficient
- Add family background controls (`parent_education`)

```{r}

```

### 6.3 Heterogeneous Effects

Ask Claude to test whether returns to `years_education` vary by:
- `gender`
- `race`
- `urban` vs rural location

```{r}

```

## 7. Media Effects with Controls

**Dataset: data/media_experiment_full.csv**

**Description**: Experimental data with extensive pre-treatment covariates.

**Variables**:

- `participant_id`: Individual identifier (int)
- `treatment`: Received partisan news treatment (lgl)
- `attitude_change`: Change in political attitudes (dbl)
- `pretreatment_attitude`: Baseline attitude (dbl)
- `political_knowledge`: Political knowledge score (int)
- `party_strength`: Strength of party ID, 1-7 (int)
- `age`: Age in years (int)
- `education`: Education level (chr)
- `news_consumption`: Regular news habits (chr)
- `political_interest`: Interest in politics, 1-10 (int)

### 7.1 Experimental Analysis with Covariates

```{r}
# Load the dataset
media_data <- read_csv("data/media_experiment_full.csv")

# Explore the data
glimpse(media_data)
```

Work with Claude to analyze the effect of `treatment` on `attitude_change` with and without covariate adjustment.

```{r}

```

### 7.2 Improving Precision

Ask Claude to:
- Show how adding covariates (`pretreatment_attitude`, `political_knowledge`, `age`, `political_interest`) as controls improves precision of the treatment effect estimate on `attitude_change`

```{r}

```

### 7.3 Specification Choices

Work with Claude to explore how different model specifications (different combinations of `pretreatment_attitude`, `political_knowledge`, `party_strength`, `age`, `education`, `news_consumption`, `political_interest`) affect the estimated treatment effect on `attitude_change`.

```{r}

```

## 8. Campaign Finance Multivariate

**Dataset: data/campaign_complete.csv**

**Description**: Comprehensive campaign finance and election data.

**Variables**:

- `candidate_id`: Candidate identifier (int)
- `vote_share`: Percentage of votes won (dbl)
- `total_spending`: Campaign spending in millions (dbl)
- `incumbent`: Incumbent status (lgl)
- `challenger_quality`: Challenger held prior office (lgl)
- `partisan_lean`: District partisan lean (dbl)
- `president_approval`: Presidential approval in district (dbl)
- `unemployment`: District unemployment rate (dbl)
- `media_markets`: Number of media markets (int)
- `primary_contested`: Had contested primary (lgl)

### 8.1 Money in Politics

```{r}
# Load the dataset
campaign_data <- read_csv("data/campaign_complete.csv")

# Look at the structure
glimpse(campaign_data)
```

### 8.2 Relative Importance

Use Claude to build a comprehensive model of `vote_share` using `total_spending`, `incumbent`, `challenger_quality`, `partisan_lean`, `president_approval`, `unemployment`, and `primary_contested`.

Work with Claude to:
- Standardize variables (`total_spending`, `partisan_lean`, `president_approval`, `unemployment`) for comparison
- Assess relative importance of `incumbent`, `total_spending`, and `partisan_lean`
- Test for interactions between `incumbent` and `total_spending`, and between `partisan_lean` and `president_approval`

```{r}

```

# Segment B — Categorical Predictors and Interactions

# Non-AI Exercises

## 1. Categorical Variables in Regression

### 1.1 Multiple Choice: Dummy Variables

When including a categorical variable with 4 categories in regression, how many dummy variables do we create?

a) 1
b) 2
c) 3
d) 4

Answer: _____

### 1.2 Fill in the Blanks: Reference Categories

For categorical variables in regression:

1. The excluded category is called the _____ category
2. Coefficients represent differences from the _____
3. With k categories, we include _____ dummy variables
4. The choice of reference category affects _____, not _____
5. R automatically creates dummies for _____ variables

Word bank: reference, reference category, k-1, interpretation, model fit, factor

### 1.3 True or False: Categorical Predictors

Mark each statement as True (T) or False (F):

_____ We must convert text categories to numbers before regression  
_____ The reference category has a coefficient of zero  
_____ Factor levels should be meaningful, not arbitrary  
_____ We can include all k dummies if we remove the intercept  

## 2. Interpreting Categorical Coefficients

### 2.1 Code Detective: Factor Regression

Given this output, what does the coefficient 5.2 mean?

```r
lm(salary ~ factor(department), data = company)
# Coefficients:
# (Intercept)                    45.0
# factor(department)Engineering   5.2
# factor(department)Marketing    -2.1
# factor(department)Sales        -0.5
```

The 5.2 means: _____

### 2.2 Match: Interpretation

Match coefficients with their interpretations:

**Coefficients:**

a) (Intercept) = 50
b) genderFemale = -5
c) partyRepublican = 10
d) educationGrad = 15

**Interpretations:**

1. Republicans score 10 points higher than Democrats
2. The reference group's average is 50
3. Graduates score 15 points higher than the reference
4. Females score 5 points lower than males

Matches: a = _____, b = _____, c = _____, d = _____

### 2.3 Spot the Error: Dummy Trap

Assuming people in this dataset are only male or female, what's wrong with this model specification?

```r
lm(outcome ~ male + female + age, data = data)
```

Problem: _____

## 3. Interaction Terms

### 3.1 Multiple Choice: Interactions

An interaction term between X₁ and X₂ tests whether:

a) X₁ and X₂ are correlated
b) The effect of X₁ depends on X₂
c) X₁ causes X₂
d) X₁ and X₂ have similar effects

Answer: _____

### 3.2 Fill in the Code: Creating Interactions

Complete this code to add an interaction:

```r
# Main effects only
model1 <- lm(y ~ x1 + x2, data = data)

# With interaction (x2)
model2 <- lm(y ~ x1 * _____, data = data)  # Shorthand

```

### 3.3 True or False: Interaction Effects

Mark each statement as True (T) or False (F):

_____ X₁*X₂ automatically includes main effects in R  
_____ Interactions must be between continuous variables  
_____ The main effect of X₁ is its effect when X₂ = 0  
_____ Centering variables changes interaction interpretation  
_____ Significant interactions mean main effects don't matter  

## 4. Interpreting Interactions

### 4.1 Code Detective: Interaction Meaning

Given: Y = 10 + 2X₁ + 3X₂ - 0.5X₁*X₂

What is the effect of X₁ when X₂ = 4?

Effect = _____

### 4.2 Match: Interaction Types

Match interaction types with examples:

**Types:**

a) Continuous × Continuous
b) Continuous × Categorical
c) Categorical × Categorical

**Examples:**

1. Income effect varies by gender
2. Age and experience interact for salary
3. Party and region interact for voting

Matches: a = _____, b = _____, c = _____

### 4.3 Multiple Choice: Marginal Effects

In a model with interactions, the marginal effect of X₁:

a) Is constant across all values of X₂
b) Depends on the value of X₂
c) Equals the main effect coefficient
d) Cannot be calculated

Answer: _____

# AI Exercises

**Tips for Working with Claude:**

- Ask for **R code using only tidyverse** (no other packages)
- Request **simple, focused answers** to your specific question—not complex analyses
- Ask Claude to **explain what the code is doing** since you're learning
- Avoid asking for visualizations or plots in these exercises
- Include the output of `glimpse()` in your prompt so Claude knows your variable names

**Example prompt:** "Using tidyverse in R, run a regression with lm() predicting salary from gender and years_experience, including an interaction term (gender * years_experience). Explain how to interpret the interaction coefficient. Keep the code simple and explain what each line does. Here is what my data looks like: [paste glimpse output]"

For each AI exercise:
- Work with Claude to analyze the data
- Record your prompts and key findings

## 5. Party and Demographics

**Dataset: data/party_demographics.csv**

**Description**: Voter registration data with party affiliation and demographics.

**Variables**:

- `voter_id`: Unique identifier (int)
- `party`: Democrat, Republican, Independent, Other (chr)
- `age_group`: 18-29, 30-44, 45-64, 65+ (chr)
- `education`: HS, Some College, College, Graduate (chr)
- `income_bracket`: <30k, 30-60k, 60-100k, >100k (chr)
- `race`: White, Black, Hispanic, Asian, Other (chr)
- `political_engagement`: Engagement score 0-100 (dbl)
- `trust_institutions`: Institutional trust 0-100 (dbl)
- `voted_2020`: Voted in 2020 election (lgl)
- `ideology_score`: Liberal-conservative 1-7 (dbl)

### 5.1 Categorical Predictors

```{r}
# Load the dataset
party_data <- read_csv("data/party_demographics.csv")

# Examine the structure
glimpse(party_data)
```

### 5.2 Reference Category Effects

Work with Claude to model political engagement using categorical predictors.

Ask Claude to help you:
- Change reference category for age_group

```{r}

```

## 6. Gender and Career Outcomes

**Dataset: data/career_gender_interact.csv**

**Description**: Professional career data examining gender effects.

**Variables**:

- `employee_id`: Unique identifier (int)
- `salary`: Annual salary in thousands (dbl)
- `gender`: Male, Female, Non-binary (chr)
- `years_experience`: Years of experience (int)
- `department`: Engineering, Sales, Marketing, HR, Finance (chr)
- `performance_rating`: Annual rating 1-5 (dbl)
- `hours_worked`: Average weekly hours (dbl)
- `has_mentor`: Has workplace mentor (lgl)
- `leadership_role`: In leadership position (lgl)
- `education_level`: Highest degree (chr)

### 6.1 Gender Main Effects

```{r}
# Load the dataset
career_data <- read_csv("data/career_gender_interact.csv")

# Check the structure
glimpse(career_data)
```

### 6.2 Testing for Interactions

Use Claude to analyze gender salary gaps with proper controls.

Work with Claude to:
- Test if experience returns differ by gender
- Examine department × gender interactions

```{r}

```

## 7. Regional Political Differences

**Dataset: data/regional_politics.csv**

**Description**: State-level political outcomes by region.

**Variables**:

- `state`: State name (chr)
- `region`: Northeast, South, Midwest, West (chr)
- `subregion`: More detailed regional classification (chr)
- `dem_vote_share`: Democratic presidential vote share (dbl)
- `turnout`: Voter turnout percentage (dbl)
- `median_income`: State median income (dbl)
- `urban_percent`: Percent urban population (dbl)
- `college_percent`: Percent with college degree (dbl)
- `evangelical_percent`: Percent evangelical (dbl)
- `union_membership`: Union membership rate (dbl)

### 7.1 Regional Categories

```{r}
# Load the dataset
regional_data <- read_csv("data/regional_politics.csv")

# Explore the data
glimpse(regional_data)
```

### 7.2  Region × Demographics

Work with Claude to test whether demographic effects vary by region.

```{r}

```

## 8. Policy Implementation Effects

**Dataset: data/policy_heterogeneity.csv**

**Description**: Policy effects across different contexts.

**Variables**:

- `unit_id`: Geographic unit identifier (int)
- `policy_implemented`: Policy adopted (lgl)
- `outcome`: Policy outcome measure (dbl)
- `implementation_quality`: High, Medium, Low (chr)
- `prior_policies`: Number of related policies (int)
- `state_capacity`: Administrative capacity index (dbl)
- `political_support`: Support for policy 0-100 (dbl)
- `urban_rural`: Urban, Suburban, Rural (chr)
- `region`: Geographic region (chr)
- `years_since_implementation`: Time since adoption (int)

### 8.1 Implementation Categories

```{r}
# Load the dataset
policy_data <- read_csv("data/policy_heterogeneity.csv")

# Look at the structure
glimpse(policy_data)
```

### 8.2 Context-Dependent Effects

Use Claude to analyze how implementation quality affects outcomes.

Work with Claude to:
- Test urban/rural differences in policy effects
- Examine quality × support interactions

```{r}

```
