---
title: "Week 9 · Class 1: Practice Exercises"
subtitle: "Statistical Tests & Confidence Intervals"
date: "2026"
format:
  pdf:
    toc: false
    number-sections: true
    colorlinks: true
---

```{r}
#| include: false
library(tidyverse)
library(knitr)
library(formatR)
library(broom)
# When running chunks interactively, set the working directory to this file's
# folder so the relative data/ paths resolve. (During render, knitr already
# uses the document's directory, so this is skipped.)
if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable()) {
  .doc_path <- rstudioapi::getSourceEditorContext()$path
  if (length(.doc_path) == 1 && nzchar(.doc_path)) setwd(dirname(.doc_path))
}
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = FALSE)
```

# Segment A — Statistical Tests

# Non-AI Exercises

## 1. Understanding Standard Error

### 1.1 Conceptual Questions

**1. What is standard error?**

a) The standard deviation of your sample data
b) The standard deviation of the sampling distribution
c) The average error in your measurements
d) The difference between sample mean and population mean

Answer: _____

**2. What happens to standard error when you double your sample size?**

a) It doubles
b) It stays the same
c) It decreases by √2 (approximately 1.41)
d) It is cut in half

Answer: _____

**3. Which factor does NOT affect the standard error?**

a) Sample size
b) Population mean
c) Sample standard deviation
d) None of the above (all affect SE)

Answer: _____

### 1.2 Calculating Standard Error

Given a sample with n = 36 and s = 12:

Calculate SE = s / √n

SE = _____ / _____ = _____

### 1.3 True or False: Standard Error Concepts

Mark each statement as True (T) or False (F):

_____ Larger samples always have smaller standard errors  
_____ Standard error measures the precision of your estimate  
_____ To halve the SE, you need 4× the sample size  
_____ SE depends on both sample size and variability  
_____ Standard error is the same as standard deviation  

## 2. The t-Distribution

### 2.1 Multiple Choice: t-Distribution Basics

**1. The t-distribution was developed to solve what problem?**

a) Computing means of large samples
b) Dealing with uncertainty when estimating standard deviation from small samples
c) Testing multiple hypotheses at once
d) Creating confidence intervals

Answer: _____

**2. As sample size increases, the t-distribution:**

a) Gets wider tails
b) Approaches the normal distribution
c) Becomes bimodal
d) Stays exactly the same

Answer: _____

**3. For a sample of n = 10, the degrees of freedom are:**

a) 10
b) 9
c) 11
d) 5

Answer: _____

### 2.2 Conceptual Understanding

**Why does the t-distribution have heavier tails than the normal distribution?**

Explain in 1-2 sentences:

_____________________________________________________________________

_____________________________________________________________________

## 3. One-Sample t-Tests

### 3.1 Hypothesis Testing Setup

A pollster wants to know whether presidential approval differs from 50%.

**Null Hypothesis (H₀)**: _______________________________________________

**Alternative Hypothesis (H₁)**: _________________________________________

**Test statistic formula**: t = _________________________________________

### 3.2 Practice Calculation

A poll of n = 25 voters finds a mean approval of 47.5% with s = 10. The critical value at α = 0.05 with df = 24 is 2.064.

**Step 1: Calculate SE**

SE = s / √n = _____ / _____ = _____

**Step 2: Calculate the t-statistic**

t = (x̄ − μ₀) / SE = (_____ − _____) / _____ = _____

**Step 3: Conclusion**

Is |t| > 2.064? _____

Conclusion: _________________________________________________________

### 3.3 Interpreting p-values

**If p = 0.08, what does this mean?**

a) There's an 8% chance the null hypothesis is true
b) If the null is true, there's an 8% chance of seeing this result or more extreme
c) The effect size is 8%
d) We should reject the null hypothesis

Answer: _____

## 4. Two-Sample t-Tests

### 4.1 Match: Research Questions to Comparisons

Match each research question to what it's comparing:

**Research Questions:**

a) Do Democrats and Republicans differ in political knowledge?  
b) Does a new curriculum improve test scores?  
c) Are battleground states more competitive?  
d) Do younger voters prefer different candidates?  

**Comparisons:**

1. Treatment vs. Control groups  
2. Two different age groups  
3. Two political parties  
4. Battleground vs. Safe states  

Matches: a = _____, b = _____, c = _____, d = _____

### 4.2 Understanding the SE of Differences

The SE of the difference between two means is:

$$SE_{diff} = \sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}$$

**True or False:**

_____ The SE of difference equals the average of the two group SEs
_____ Larger samples lead to smaller SE of difference
_____ We add variances, not standard errors
_____ Equal sample sizes minimize the SE of difference

### 4.3 Practice Two-Sample Test

Two groups of students take a civic knowledge test:

- **Group 1** (took a government course): n₁ = 40, x̄₁ = 72, s₁ = 12
- **Group 2** (did not): n₂ = 40, x̄₂ = 65, s₂ = 15

Use a critical value of 2.00 (α = 0.05).

**Step 1: Calculate the difference in means**

Difference = _____ − _____ = _____

**Step 2: Calculate the SE of the difference**

SE~diff~ = √(s₁²/n₁ + s₂²/n₂) = √(_____/_____ + _____/_____) = _____

**Step 3: Calculate the t-statistic**

t = Difference / SE~diff~ = _____ / _____ = _____

**Step 4: Interpret**

Significant? _____

Why? _______________________________________________________________

### 4.4 When t-Tests Can Go Wrong

**1. Which situation is most likely to make a two-sample t-test unreliable?**

a) The two groups have slightly different variances
b) n = 12 per group and the outcome is strongly skewed
c) n = 200 per group and the outcome is not perfectly normal
d) The outcome is measured on a 0-100 scale

Answer: _____

**2. R's `t.test()` applies Welch's correction by default. What problem does this address?**

a) Outliers in the outcome variable
b) Unequal variances between the two groups
c) Non-random sampling
d) Small sample sizes

Answer: _____

**3. True or False:**

_____ With n > 30 per group, the t-test usually performs well even when the data are skewed
_____ Extreme outliers can distort both the means and the variances a t-test relies on
_____ The classical (non-Welch) t-test assumes the two groups have equal variances
_____ A t-test requires the outcome to be perfectly normally distributed in every case

## 5. Standard Error in Regression

### 5.1 Multiple Choice: Regression SE

**1. The standard error of a regression coefficient tells us:**

a) How much the coefficient would vary across repeated samples
b) The average residual
c) The strength of the relationship
d) The probability the coefficient is zero

Answer: _____

**2. Which increases the SE of a regression coefficient?**

a) Larger sample size
b) More variability in the outcome (larger residuals)
c) More variability in the predictor
d) Adding more control variables

Answer: _____

**3. A coefficient is 0.08 with SE = 0.02. Is it significant at α = 0.05?**

a) Yes, because |0.08 / 0.02| = 4 > 1.96
b) No, because 0.08 is small
c) Cannot determine without p-value
d) Yes, because SE is small

Answer: _____

### 5.2 Reading Regression Output

Interpret this regression output:

```
               Estimate  Std.Error  t.value  p.value
(Intercept)    45.200    3.100      14.58    <0.001
ad_spending     0.035    0.008       4.38    <0.001
incumbent       8.500    1.200       7.08    <0.001
```

**Questions:**

1. For every $1,000 increase in ad_spending, vote share increases by: _____

2. The incumbent coefficient SE is 1.200. What does this mean?

   _________________________________________________________________

3. Which predictor has the most precise estimate (smallest SE relative to estimate)?

   _________________________________________________________________

### 5.3 Coefficient Plot Interpretation

Consider these three coefficients with 95% confidence intervals:

- **Education**: 0.50 ± 0.20 (CI: 0.30 to 0.70)
- **Age**: 0.15 ± 0.18 (CI: -0.03 to 0.33)
- **Income**: -0.05 ± 0.25 (CI: -0.30 to 0.20)

**Questions:**

1. Which coefficients are statistically significant? _____

2. How can you tell from the confidence intervals? _____

3. Which estimate is most precise? _____ Why? _____

## 6. t-Tests in Regression Context

### 6.1 True or False: Understanding Regression t-Tests

_____ Every regression coefficient is automatically tested against zero
_____ The t-statistic is calculated as: coefficient / SE
_____ In multiple regression, t-tests are conditional on other variables
_____ A two-sample t-test is equivalent to regression with a binary predictor
_____ Larger t-statistics indicate weaker evidence

### 6.2 Conditional Tests

**Simple regression**: vote_share ~ campaign_spending
- Coefficient: 0.045, SE: 0.010, t: 4.50

**Multiple regression**: vote_share ~ campaign_spending + incumbent
- Coefficient: 0.035, SE: 0.008, t: 4.38

**Questions:**

1. Why did the coefficient change from 0.045 to 0.035?

   _________________________________________________________________

2. What does the t-test mean in the simple regression?

   _________________________________________________________________

3. What does the t-test mean in the multiple regression?

   _________________________________________________________________

### 6.3 Match: Interpretation of t-Statistics

Match each t-value with its interpretation (α = 0.05, large sample):

**t-values:**
a) t = 1.2
b) t = 2.5
c) t = 0.4
d) t = -3.8

**Interpretations:**
1. Strong evidence against null (significant)
2. Moderate evidence against null (significant)
3. Weak evidence, fail to reject null
4. Very weak evidence, fail to reject null

Matches: a = _____, b = _____, c = _____, d = _____

## 7. Degrees of Freedom

### 7.1 Calculating Degrees of Freedom

1. One-sample t-test with n = 50: df = _____

2. Two-sample t-test with n₁ = 30, n₂ = 40: df ≈ _____

3. Regression with n = 100 and 3 predictors: df = _____

### 7.2 Multiple Choice: DF Concepts

**Why do we subtract 1 for degrees of freedom in a one-sample test?**

a) To make the test more conservative, by convention
b) Because we use up one degree of freedom estimating the mean
c) Because the first observation carries no information
d) Because sample sizes are always off by one

Answer: _____

# AI Exercises

## Worked Example: What an AI-Assisted Solution Should Look Like

Before you start, here is a complete worked solution in the format we expect from the exercises below. Use this as a template for the depth and style of your write-up.

**Question:** Does the GOTV canvassing treatment increase turnout?

**Prompt to Claude:** "Using tidyverse in R, load `data/gotv_experiment.csv`, compute mean turnout by treatment group, and run a two-sample t-test of turnout by treatment. Briefly explain the output."

**R code returned:**

```{r}
#| eval: false
gotv <- read_csv("data/gotv_experiment.csv")

gotv %>%
  group_by(treatment) %>%
  summarise(mean_turnout = mean(turnout), n = n())

t.test(turnout ~ treatment, data = gotv)
```

**Sample interpretation (what you write):**

> The canvassed group turned out at 48.8% vs. 44.4% for the control group --- a 4.4 percentage point difference (t = 0.99, df = 498, p = 0.325). Because p > 0.05, we fail to reject the null of no difference: a gap this size could easily arise by chance in a study of 500 voters. The 95% CI for the difference runs from -4.4 to +13.2 points, so it includes zero. Note what this does *not* say. The interval also includes effects as large as 13 points, which would be substantively enormous for a GOTV campaign. This study simply lacks the precision to distinguish "no effect" from "a large effect," so the honest conclusion is that we cannot tell --- not that canvassing does nothing.

Use this structure --- (1) prompt, (2) code, (3) numerical result, (4) one-paragraph interpretation that names the test, the p-value, and the substantive conclusion --- for the exercises that follow.

**Tips for Working with Claude:**

- Ask for **R code using only tidyverse** (no other packages)
- Request **simple, focused answers** to your specific question—not complex analyses
- Ask Claude to **explain what the code is doing** since you're learning
- Avoid asking for visualizations or plots in these exercises
- Include the output of `glimpse()` in your prompt so Claude knows your variable names

**Example prompt:** "Using tidyverse in R, run a t.test() comparing turnout between the Control and Canvassed groups. Explain what the t-statistic, p-value, and standard error mean. Keep the code simple and explain what each line does. Here is what my data looks like: [paste glimpse output]"

For each exercise, work with Claude to conduct analyses and interpret results.

**Datasets Used in This Section:**

1. **data/gotv_experiment.csv** (n=500): GOTV experiment with treatment/control groups
2. **data/political_knowledge.csv** (n=200): News consumption and political knowledge
3. **data/campaign_spending.csv** (n=200): Campaign spending, incumbency, and vote share

## 8. Get-Out-The-Vote Analysis

**Dataset: data/gotv_experiment.csv**

GOTV experiment data to practice two-sample t-tests.

```{r}
# Load the dataset
gotv_data <- read_csv("data/gotv_experiment.csv")

# Examine the data
glimpse(gotv_data)
```

**Variables**:

- `voter_id`: Unique voter identifier (int, 1-500)
- `treatment`: Experimental condition (chr: "Control" or "Canvassed")
- `turnout`: Whether voter turned out (int: 0 or 1)
- `turnout_pct`: Turnout as percentage (dbl: 0 or 100)

**Work with Claude to:**

1. Calculate summary statistics by group
2. Conduct a two-sample t-test

```{r}

```

## 9. Political Knowledge Study

**Dataset: data/political_knowledge.csv**

Relationship between news consumption and political knowledge.

```{r}
# Load the dataset
knowledge_data <- read_csv("data/political_knowledge.csv")

# Examine the data
glimpse(knowledge_data)
```

**Variables**:

- `person_id`: Unique person identifier (int, 1-200)
- `watches_news`: Watches news regularly (chr: "Yes" or "No")
- `knowledge_score`: Political knowledge score (dbl: 0-10 scale)

**Ask Claude to help:**

1. Test if news watchers have higher knowledge
2. Calculate and interpret the standard error
3. Discuss whether this is causal evidence

```{r}

```

## 10. Campaign Spending Regression

**Dataset: data/campaign_spending.csv**

Relationship between campaign spending and electoral outcomes. This is the same dataset used in lecture.

```{r}
# Load the dataset
campaign_data <- read_csv("data/campaign_spending.csv")

# Examine the data
glimpse(campaign_data)
```

**Variables**:

- `candidate_id`: Unique candidate identifier (int, 1-200)
- `campaign_spending`: Campaign spending in thousands of dollars (dbl)
- `incumbent`: Incumbent status (int: 0 = challenger, 1 = incumbent)
- `vote_share`: Vote share percentage (dbl: 0-100)

**Work with Claude to:**

1. Run a simple regression of vote_share on campaign_spending
2. Run a multiple regression adding incumbent status
3. Interpret all t-statistics
4. Explain how the spending coefficient changed

```{r}

```

# Segment B — Confidence Intervals

# Non-AI Exercises

## 1. Understanding Confidence Intervals

### 1.1 Conceptual Understanding

**1. What does a 95% confidence interval tell us?**

a) There's a 95% chance the true parameter is in this interval
b) 95% of the data falls within this interval
c) If we repeated the study 100 times, about 95 CIs would contain the true parameter
d) The parameter varies 95% of the time

Answer: _____

**2. A 95% CI for a coefficient is [2.1, 5.8]. What can we conclude?**

a) The effect is not statistically significant at α = 0.05
b) The effect is statistically significant at α = 0.05
c) We need more information to determine significance
d) The true value is exactly 3.95 (the midpoint)

Answer: _____

**3. Which statement about confidence intervals is TRUE?**

a) Wider intervals indicate more precision
b) Higher confidence levels require narrower intervals
c) Larger samples generally produce narrower CIs
d) CIs and p-values provide completely different information

Answer: _____

### 1.2 True or False: CI Concepts

Mark each statement as True (T) or False (F):

_____ A 99% CI is wider than a 95% CI for the same data
_____ If a 95% CI excludes zero, then p < 0.05
_____ The confidence interval is random, but the parameter is fixed
_____ Narrower CIs always mean larger effect sizes
_____ Sample size affects the width of confidence intervals

## 2. Calculating Confidence Intervals

### 2.1 Manual CI Calculation

Given the following information from a regression:

- Coefficient (β̂) = 0.45
- Standard Error (SE) = 0.12
- Sample size (n) = 60, so df = 58
- Critical value for 95% CI with df = 58: t~critical~ ≈ 2.00

**Step 1: Calculate the margin of error**

Margin = t~critical~ × SE = _____ × _____ = _____

**Step 2: Calculate CI bounds**

Lower bound = β̂ - Margin = _____ - _____ = _____

Upper bound = β̂ + Margin = _____ + _____ = _____

**95% CI**: [_____, _____]

### 2.2 Interpreting Different Confidence Levels

For the same coefficient (β̂ = 0.45, SE = 0.12, df = 58):

Calculate confidence intervals at different levels:

**90% CI** (t~critical~ ≈ 1.67):

Lower = _____, Upper = _____, Width = _____

**95% CI** (t~critical~ ≈ 2.00):

Lower = _____, Upper = _____, Width = _____

**99% CI** (t~critical~ ≈ 2.66):

Lower = _____, Upper = _____, Width = _____

**Question**: What happens to the width as confidence increases? _____

### 2.3 CI and Significance

For each confidence interval, determine if the coefficient is significant at α = 0.05:

**1. 95% CI: [0.2, 0.8]**

Significant? _____ Why? _____________________________

**2. 95% CI: [-0.1, 0.5]**

Significant? _____ Why? _____________________________

**3. 95% CI: [-2.5, -0.8]**

Significant? _____ Why? _____________________________

## 3. Three Roads to the Same Conclusion

### 3.1 Connecting SE, t-statistic, p-value, and CI

Given regression output:

```
Coefficient: 0.035
Std. Error: 0.008
t-value: ?
p-value: ?
```

**Step 1: Calculate t-statistic**

t = Coefficient / SE = _____ / _____ = _____

**Step 2: With df = 100, is |t| > 1.98?**

_____ (Yes/No), so the coefficient is _____ (significant/not significant)

**Step 3: Calculate 95% CI (t~critical~ ≈ 1.98)**

CI = 0.035 ± (1.98 × 0.008) = [_____, _____]

**Step 4: Does the CI exclude zero?**

_____ (Yes/No), confirming the coefficient is _____ (significant/not significant)

### 3.2 Comparing Two Studies

**Study A** (small sample, n = 40):

- β̂ = 0.75, SE = 0.30, t = 2.50, p = 0.017
- 95% CI: [0.14, 1.36]

**Study B** (large sample, n = 500):

- β̂ = 0.78, SE = 0.09, t = 8.67, p < 0.001
- 95% CI: [0.60, 0.96]

**Questions:**

1. Which study has a more precise estimate? _____

2. Why? (Explain in terms of SE or CI width)

   _________________________________________________________________

3. Both studies found similar effect sizes. Why does Study B have a smaller p-value?

   _________________________________________________________________

### 3.3 Match: Statistical Concepts

Match each concept to its interpretation:

**Concepts:**
a) Standard Error = 0.15
b) t-statistic = 3.2
c) p-value = 0.002
d) 95% CI = [0.25, 0.65]

**Interpretations:**
1. Coefficient is 3.2 standard errors from zero
2. Range of plausible values for the parameter
3. Measure of sampling variability/precision
4. Probability of seeing this result if null is true

Matches: a = _____, b = _____, c = _____, d = _____

## 4. Sample Size and Precision

### 4.1 True or False: Sample Size Effects

Mark each statement as True (T) or False (F):

_____ Doubling sample size doubles precision (halves SE)
_____ Larger samples produce narrower confidence intervals
_____ With n = 10,000, even tiny effects can be statistically significant
_____ Sample size affects the point estimate (coefficient)
_____ Critical values get smaller as sample size increases

### 4.2 Practical Implications

A pilot study (n = 30) finds:
- β̂ = 5.2, SE = 2.5, p = 0.048
- 95% CI: [0.1, 10.3]

**Questions:**

1. Is this result statistically significant at α = 0.05? _____

2. Is the confidence interval informative? Why or why not?

   _________________________________________________________________

3. What would you recommend for a follow-up study?

   _________________________________________________________________

## 5. Writing About Confidence Intervals

### 5.1 Identify Good vs. Bad Writing

Rate each write-up as Good (G) or Bad (B):

**Write-up A:**
"The coefficient for education was 2.5 (95% CI: [1.2, 3.8], p = 0.001)."

Rating: _____ Problem (if bad): _________________________________

**Write-up B:**
"Education significantly predicts income (β = 2.5, 95% CI: [1.2, 3.8], t(148) = 3.85, p = 0.001). Each additional year of education is associated with $2,500 higher annual income. This effect is both statistically and practically meaningful."

Rating: _____ Strengths: _____________________________________

**Write-up C:**
"There is a 95% probability that the true effect of education is between 1.2 and 3.8."

Rating: _____ Problem (if bad): _________________________________

**Write-up D:**
"Education had no effect (p = 0.12)."

Rating: _____ Problem (if bad): _________________________________

### 5.2 Fix the Bad Write-Ups

Rewrite one of the bad examples from 5.1 to make it good:

Original: _________________________________________________________

Improved: _________________________________________________________

_________________________________________________________________

_________________________________________________________________

## 6. Critical Values and qt()

### 6.1 Confidence Level Practice

Match each confidence level to the probability used in qt():

**Confidence Levels:**
a) 90% CI
b) 95% CI
c) 99% CI
d) 99.9% CI

**qt() probabilities:**
1. 0.995
2. 0.975
3. 0.95
4. 0.9995

Matches: a = _____, b = _____, c = _____, d = _____

**Why these values?** (Explain two-tailed logic)

_________________________________________________________________

_________________________________________________________________

## 7. Common Pitfalls

### 7.1 Statistical vs. Practical Significance

**Scenario**: A study with n = 50,000 finds:

- β̂ = 0.08, SE = 0.02, p < 0.001
- 95% CI: [0.04, 0.12]
- Outcome is measured 0-100 scale

**Questions:**

1. Is this statistically significant? _____

2. Is this practically significant? Why or why not?

   _________________________________________________________________

3. What's the problem with focusing only on the p-value?

   _________________________________________________________________

### 7.2 Non-Significant ≠ No Effect

**Study results**: β̂ = 6.5, SE = 4.2, p = 0.13
- 95% CI: [-2.0, 15.0]

**Bad interpretation**: "There is no effect (p > 0.05)."

**Questions:**

1. What's wrong with this interpretation?

   _________________________________________________________________

2. What does the CI tell us?

   _________________________________________________________________

3. Write a better interpretation:

   _________________________________________________________________

### 7.3 Multiple Testing

A researcher tests 20 unrelated predictors of turnout. None of them has any true relationship with turnout. She uses α = 0.05 for every test.

**1. About how many of the 20 tests would you expect to come out "significant" purely by chance?**

a) 0
b) 1
c) 5
d) 20

Answer: _____

**2. She reports only the one significant predictor and does not mention the other 19 tests. What is the problem?**

a) Nothing — she reported a result that met the standard threshold
b) The reported p-value is misleading, because it ignores the 19 other chances she gave herself to find something
c) She should have collected a larger sample
d) She should have used a 99% confidence interval instead

Answer: _____

**3. Name one thing she could do to report this honestly:**

_________________________________________________________________

### 7.4 Match: Pitfalls to Solutions

Match each pitfall to its solution:

**Pitfalls:**
a) Confusing statistical with practical significance
b) Overinterpreting non-significant results
c) Ignoring confidence interval width
d) Assuming causation from correlation

**Solutions:**
1. Always report and discuss effect sizes
2. Check for confounders; need experimental design
3. Acknowledge uncertainty; consider power
4. Comment on CI width/precision

Matches: a = _____, b = _____, c = _____, d = _____

# AI Exercises

**Tips for Working with Claude:**

- Ask for **R code using only tidyverse** (no other packages)
- Request **simple, focused answers** to your specific question—not complex analyses
- Ask Claude to **explain what the code is doing** since you're learning
- Avoid asking for visualizations or plots in these exercises
- Include the output of `glimpse()` in your prompt so Claude knows your variable names

**Example prompt:** "Using tidyverse in R, run a regression of test_score on spending_per_student, then use confint() to get the 95% confidence interval. Explain what the confidence interval tells us about the spending effect. Keep the code simple and explain what each line does. Here is what my data looks like: [paste glimpse output]"

## 8. Education and Test Scores

**Dataset: data/education_spending.csv**

Education spending and test score data to practice confidence intervals.

```{r}
# Load the dataset
education_data <- read_csv("data/education_spending.csv")

# Examine the data
glimpse(education_data)
```

**Variables**:

- `district_id`: Unique district identifier (int, 1-100)
- `spending_per_student`: Per-student spending in dollars (dbl)
- `test_score`: Average district test score (dbl)

**Work with Claude to:**

1. Run the regression and examine output
2. Calculate 95% confidence intervals using `confint()`
3. Interpret the confidence interval for spending
4. Discuss whether the interval is narrow or wide
5. Write a complete results section

```{r}

```

## 9. Campaign Spending Analysis

**Dataset: data/campaign_spending.csv**

Campaign spending with multiple predictors.

```{r}
# Load the dataset
campaign_data <- read_csv("data/campaign_spending.csv")

# Examine the data
glimpse(campaign_data)

```

**Variables**:

- `candidate_id`: Unique candidate identifier (int, 1-200)
- `campaign_spending`: Campaign spending in thousands of dollars (dbl)
- `incumbent`: Incumbent status (int: 0 = challenger, 1 = incumbent)
- `vote_share`: Vote share percentage (dbl)

**Ask Claude to help:**

1. Get confidence intervals for all coefficients
2. Interpret each CI substantively
3. Write up results following best practices

```{r}

```

## 10. Comparing Sample Sizes

**Datasets: data/voter_turnout_small.csv & data/voter_turnout_large.csv**

Demonstrate the effect of sample size on CI width.

```{r}
# Load small sample dataset
voter_study_small <- read_csv("data/voter_turnout_small.csv")
glimpse(voter_study_small)

# Load large sample dataset
voter_study_large <- read_csv("data/voter_turnout_large.csv")
glimpse(voter_study_large)

```

**Variables** (both datasets):

- `precinct_id`: Unique precinct identifier (int)
- `canvassing_contacts`: Number of canvassing contacts (int)
- `turnout_pct`: Voter turnout percentage (dbl)

**Work with Claude to:**

1. Compare the point estimates from both models
2. Compare the standard errors
3. Compare the 95% confidence intervals

```{r}

```

## 11. Multiple Confidence Levels

**Dataset: data/public_opinion.csv**

The same model reported at three different confidence levels.

```{r}
# Load the dataset
opinion_data <- read_csv("data/public_opinion.csv")

# Examine the data
glimpse(opinion_data)
```

**Variables**:

- `respondent_id`: Unique respondent identifier (int, 1-150)
- `age`: Age in years (dbl)
- `education_years`: Years of education (dbl)
- `policy_support`: Policy support score (dbl: 0-100 scale)

**Work with Claude to:**

1. Fit `policy_support ~ age + education_years`
2. Get the 90%, 95%, and 99% CI for the `age` coefficient using `confint(..., level = )`
3. Explain what happens to the width as the confidence level rises, and why
4. Decide which level you would report, and justify it

```{r}

```

## 12. Writing Practice

Using the education model from Exercise 8, write a complete results paragraph.

**Work with Claude to:**

1. Pull the coefficient, CI, t-statistic, and p-value into one place
2. Draft a paragraph that follows the six-step template: direction and significance, statistics, effect size, precision, practical meaning, context
3. Check your draft against the "what NOT to do" list — no causal language, no "95% probability" phrasing, no numbers without interpretation

```{r}

```

## 13. Simulation: Understanding "95% Confident"

No dataset — you will simulate one.

**Work with Claude to:**

1. Write a function that draws one sample, fits a regression, and returns the 95% CI for the slope
2. Run it 100 times against a known true slope
3. Count how many of the 100 intervals contain the true value
4. Explain how this simulation illustrates what "95% confident" actually means

```{r}

```

## 14. Critical Values Practice

No dataset — this one is about `qt()`.

**Work with Claude to:**

1. Build a table of critical values for n = 10, 30, 100, and 1000 at the 90%, 95%, and 99% levels
2. Explain why the probability passed to `qt()` is (1 + level) / 2 rather than the level itself
3. Describe what happens to the critical value as df grows, and what it converges to

```{r}

```

## 15. Common Pitfalls Investigation

**Dataset: data/large_sample_study.csv**

A very large study of a very small effect.

```{r}
# Load the dataset
large_sample <- read_csv("data/large_sample_study.csv")

# Examine the data
glimpse(large_sample)
```

**Variables**:

- `person_id`: Unique person identifier (int, 1-10000)
- `intervention`: Treatment assignment (dbl: 0 = control, 1 = treated)
- `outcome`: Outcome measure (dbl: 0-100 scale)

**Work with Claude to:**

1. Fit `outcome ~ intervention` and get the coefficient, p-value, and CI
2. Express the effect as a percentage change from the control-group baseline
3. Explain how a result can be highly significant and still not matter

```{r}

```

## 16. Complete Analysis Project

Choose any dataset from this exercise and produce a complete analysis.

**Your write-up should demonstrate:**

1. A clearly stated question and an appropriate model
2. A correctly calculated confidence interval
3. A clear interpretation of the coefficient
4. A comment on precision (CI width)
5. Discussion of both statistical and practical significance
6. A write-up that follows the reporting template

```{r}

```

\newpage
